Density-dependent selection in evolutionary genetics: a lottery model of Grime's triangle

### 2016-10-31 manuscript.tex
latest

In ideal environments, we predict evolution for higher $b$ and $c$, but not lower $d$. In plants and corals, a key mechanism for winning territorial contests is rapidly outgrowing and ``shading out'' competitors; not surprisingly, rapid individual growth is a defining feature of the competitor trait cluster \citep{grime_1977,darling_2012}. Evolution for higher $b$ under high-density, competitive conditions may seem counter-intuitive given that $b$ is closely related to the low-density reproductive rate $r=b-d$, and the opposite outcome is expected in MacArthur's $r$/$K$ dichotomy \citep{macarthur_1962}. Neither particularly high nor low $b$ have been associated with the competitor strategy in plants and corals. However, for fishes, the analogous ``periodic'' strategy is characterized by enormous brood sizes as well as rapid development \citep{winemiller_1992,winemiller_2015}, suggesting a strategy of ensuring that many propagules actually end up contesting areas favorable for development (higher $b$). The evolution of $b$ in ideal environments will be discussed further in the Discussion.

@@

@@

Indeed, the classic lottery model was originally applied to territorial reef fishes, and consequently the $l_i\gg 1$ assumption was

>>>>

In ideal environments, we predict evolution for higher $b$ and $c$, but not lower $d$. In plants and corals, a key mechanism for winning territorial contests is rapidly outgrowing and ``shading out'' competitors; not surprisingly, rapid individual growth is a defining feature of the competitor trait cluster \citep{grime_1977,darling_2012}. Evolution for higher $b$ under high-density, competitive conditions may seem counter-intuitive. Neither particularly high nor low $b$ have been associated with the competitor strategy in plants and corals. However, for fishes, the analogous ``periodic'' strategy is characterized by enormous brood sizes as well as rapid development \citep{winemiller_1992,winemiller_2015}, suggesting a strategy of ensuring that many propagules actually end up contesting areas favorable for development (higher $b$). The evolution of $b$ in ideal environments will be discussed further in the Discussion.

@@

Unlike Grime's classic ternary plot for observed traits \citep{grime_1974}, our realization of Grime's triangle (Fig. 4) refers only to the direction of trait evolution. Whereas a ternary plot represents anti-correlation between traits relevant for success in different environmental archetypes, our triangle reflects the underlying environmental limits of survival.

We have made no appeal to, or attempted to explain, trade-offs between traits .

given that $b$ is closely related to the low-density reproductive rate $r=b-d$, and the opposite outcome is expected in MacArthur's $r$/$K$ dichotomy \citep{macarthur_1962}.

Apart from the prediction of evolution for higher $b$ in ideal environments,

@@
